### 2016-12-15 doc/paper/paper.tex
described Init phase and started Genetic one

\lstdefinelanguage{HTML5}{
  language=html,
  sensitive=true,	
  alsoletter={<>=-},	
  morecomment=[s]{<!-}{-->},
  tag=[s],
  otherkeywords={
  % General
  >,
  % Standard tags
  <!DOCTYPE,
  </html, <html, <head, <title, </title, <style, </style, <link, </head, <meta, />,
  % body
  </body, <body,
  % Divs
  </div, <div, </div>, 
  % Paragraphs
  </p, <p, </p>,
  % scripts
  </script, <script,
  % More tags...
  <canvas, /canvas>, <svg, <rect, <animateTransform, </rect>, </svg>, <video, <source, <iframe, </iframe>, </video>, <image, </image>, <header, </header, <article, </article
  },
  ndkeywords={
  % General
  =,
  % HTML attributes
  charset=, src=, id=, width=, height=, style=, type=, rel=, href=,
  % SVG attributes
  fill=, attributeName=, begin=, dur=, from=, to=, poster=, controls=, x=, y=, repeatCount=, xlink:href=,
  % properties
  margin:, padding:, background-image:, border:, top:, left:, position:, width:, height:, margin-top:, margin-bottom:, font-size:, line-height:,
  % CSS3 properties
  transform:, -moz-transform:, -webkit-transform:,
  animation:, -webkit-animation:,
  transition:,  transition-duration:, transition-property:, transition-timing-function:,
  }
}

@@

\begin{figure}[t]
  \begin{lstlisting}[style=htmlcssjs,language=JavaScript]
/*t dom */
function safeAdd() {
  var iframe = document.createElement("iframe");
  var whereto = document.getElementById("debug");
  var frameid = document.getElementById("frame");
  iframe.setAttribute("id", "frame");
  if (frameid) { 
    iframe.removeChild(frameid.childNodes[0]); |c \label{lbl:error} |c
    // frameid.parentNode.removeChild(frameid); |c \label{lbl:correction} |c
  } else { whereto.appendChild(iframe); } }  
  \end{lstlisting}
  \caption{Motivating Example}
  \label{fig..safeAdd.js}
\end{figure}

\begin{figure}[t]
  \begin{lstlisting}[style=htmlcssjs, language=HTML5]
<html>
  <body>
    <div id="frame"></div>
    <div id="debug"></div>
  </body>
</html>
  \end{lstlisting}

  \begin{lstlisting}[style=htmlcssjs, language=HTML5]
<html>
  <body>
    <div id="debug"></div>
  </body>
</html>
\end{lstlisting}  
\caption{Motivating Example}
\label{fig.safeAdd.html}
\end{figure}

\begin{figure}[t]
  \begin{lstlisting}[style=htmlcssjs,language=JavaScript]
/*t dom */
function revealAll() {
 for(var row=0; row<2; row++){
  for(var col=0; col<2; col++){
   var obj = document.getElementById('square_'+row+'_'+col);
   var spans = obj.getElementsByTagName('SPAN');
   spans[0].style.display='';
   spans[1].style.display='none';
   spans[1].style.color='#000000'; } } }
  \end{lstlisting}
  \caption{Motivating Example}
  \label{fig.revealAll.js}
\end{figure}

%% --------------------------------------------------------------------
\section{Test Generation Framework}
\label{sec.framework}
%% --------------------------------------------------------------------
The general workflow of our JavaScript test generation framework is represented as a sequence diagram in Figure~\ref{fig.framework.architect}. The diagram consists of four interacting components: \User, \Server\, \Validator and \Client, where \Server is the main one. It performs all preparatory steps (\emph{Initialization Phase}) as well as controls the data generation (\emph{Genetic Phase}). These two phases will be covered in details in Section~\ref{sub.sec.init.phase} and \ref{sub.sec.genetic.phase} respectively.

Every request sent to \Validator\footnote{\url{https://validator.w3.org/nu/}} with an HTML document as the body triggers a response containing the list of syntactic errors in that document. \Client is a \emph{Node.js}\footnote{\url{https://nodejs.org/en/}} application that is mainly responsible for the execution of the JavaScript code. It relies on the \emph{jsdom}\footnote{\url{https://github.com/tmpvar/jsdom}} library to simulate calls to native DOM.

Let us get a closer look at the sequence of interactions in Figure~\ref{fig.framework.architect}. First, \User provides JS file (\emph{file.js}) that contains the function under test (FUT) together with its type and a configuration file \emph{config}. These data are processed by the \Server in the \emph{Initialization Phase}. As an output we get a tuple with the following information about FUT: control flow graph (\emph{cfg}), list of branches (\emph{branches}), static information about the program's constants (\emph{cinfo}), instrumented version of the original program (\emph{ifun}) and the function's signature (\emph{sig}). \Server sends a \texttt{POST} request to the \Client to store \emph{ifun} and \emph{sig}. Second, \Server asks \User to select a branch(s) that has to be covered by the execution of FUT. For each branch our framework triggers the \emph{Genetic Phase (Loop)}. Inside of this phase it starts with a randomly generated population (\emph{Random Loop}), and continues to improve it until earthier perfect value is found or generation limit is reached. Each candidate in the population has to be evaluated. \Client parses the candidate according the know signature, invokes \emph{ifun} on the result, collects some information along the execution and sends it back to \Server, where the final fitness value is computed for that candidate (\emph{Evaluation Loop}). New population is the formed as joint result of the \emph{Crossover} and \emph{Mutation Loop}. At the end of the \emph{Fitness Loop} the best candidate with the best fitness value is reported to \User. After exhausting all branches \emph{Genetic Loop} terminates.

@@

    \begin{call}{user}{file.js + config}{server}{\textbf{end}}
      \begin{call}{server}{Initialize(file.js, config)}{server}{(cfg, branches, cinfo, ifun, sig)}
      \end{call}

@@

\begin{algorithm}[t!]
  \caption{Initialization Phase}
  \label{alg.init}
  \algsetup{linenosize=\tiny}
  \scriptsize
  \DontPrintSemicolon
  \SetAlgoVlined
  \SetKwInOut{Input}{Input}
  \SetKwInOut{Output}{Output}
  \SetKwInOut{Require}{Require}
  \SetKwProg{Fn}{Function}{}{}
  \Input{
    $fun.js$ is a JS file with FUT and type annotation
  }
  \Output{
    tuple with $(cfg, branches, cinfo)$
  }
  \Fn{$Initialize(fun.js)$}{

    $(ast, sig) \longleftarrow ParseFunAndSig(fun.js)$\label{alg1.parse.fun.sig}\;
    $ast' \longleftarrow TransAST(ast)$\;
    $cfg \longleftarrow BuildCFG(ast')$\;
    $branches \longleftarrow GetBranches(ast')$\;
    $cinfo \longleftarrow GetConstInfo(ast')$\;
    $ifun \longleftarrow Instrument(ast')$\;
    $Send(POST, /init, \{ifun, sig\})$\;
    \Return{$(cfg, branches, cinfo)$}
    }
\end{algorithm}

\begin{algorithm}[!t]
  \caption{Genetic Phase}
  \label{alg.gen}
  \algsetup{linenosize=\tiny}
  \scriptsize
  \DontPrintSemicolon
  \SetAlgoVlined
  \SetKwInOut{Input}{Input}
  \SetKwInOut{Output}{Output}
  \SetKwInOut{Require}{Require}
  \SetKwProg{Fn}{Function}{}{}
  \Fn{$Genetic(b_N, cfg, cinfo, config)$}{
    $(size_P, size_A, size_G, rate_C, rate_M) \longleftarrow Read(config)$\;
    $archive \longleftarrow \emptyset$\;
    $i \longleftarrow 0 $\;
    $pop \longleftarrow Random(cinfo, size_P)$\;
    \Repeat{$isPerfect(f^*) \vee i > size_G $}{
      $scoredPop \longleftarrow \emptyset$\;
      \ForAll{$p \in pop$}{
        $(tr, d_B, d_L, consts) \longleftarrow Send(POST, /genetic, \{p\})$\;
        \tcp{evaluate $ifun(p: sig)$}
        $(f, p) \longleftarrow Evaluate(b_N, cfg, tr, d_B, d_L)$\;
        $scoredPop \longleftarrow (f, p) \cup scoredPop$\;
        $cinfo \longleftarrow consts \cup cinfo$\;
      }
      $combo \longleftarrow scoredPop \cup archive$\;
      $cross \longleftarrow Crossover(combo, rate_C)$\;
      $mut \longleftarrow Mutation(combo, cinfo, rate_M)$\;
      $pop \longleftarrow cross \cup mut$\;
      $archive \longleftarrow GetBest(combo, size_A)$\;
      $(f^*, p^*) \longleftarrow head(archive)$\;
      $i \longleftarrow i + 1$\;
    }
    \Return{$(f^*, p^*)$} 
  }
\end{algorithm}

@@

@@

The heart of our test data generation framework is its genetic phase. The main workflow of the underlying genetic algorithm (GA) is quite standard~\cite{poli2008field}. First, random population is generated. Then, this population is evaluated against given fitness function. If perfectly fitted candidate is found or the budget is exhausted, the algorithm terminates and reports the best candidate so far. Otherwise it enters an iterative phase where after a selections procedure some candidates are crossed over and others are mutated. Afterwards a new population is formed and the evaluation repeats.

@@

\begin{description}
\item[\textbf{Q1}] How often \emph{validator} rejects generated HTML in case of the random and genetic (mutation and crossover) phases. This question characterizes potential performance issues that can occur in the generation.     
\end{description}

>>>>

\lstdefinelanguage{HTML5}{
  language=html,
  sensitive=true,	
  alsoletter={<>=-},	
  morecomment=[s]{<!-}{-->},
  tag=[s],
  otherkeywords={
  % General
  >,
  % Standard tags
  <!DOCTYPE,
  </html, <html, <head, <title, </title, <style, </style, <link, </head, <meta, />,
  % body
  </body, <body,
  % Divs
  </div, <div, </div>, 
  % Paragraphs
  </p, <p, </p>,
  % scripts
  </script, <script,
  % More tags...
  <canvas, /canvas>, <svg, <rect, <animateTransform, </rect>, </svg>, <video, <source, <iframe, </iframe>, </video>, <image, </image>, <header, </header, <article, </article, </span, <span, </span> 
  },
  ndkeywords={
  % General
  =,
  % HTML attributes
  charset=, src=, id=, width=, height=, style=, type=, rel=, href=,
  % SVG attributes
  fill=, attributeName=, begin=, dur=, from=, to=, poster=, controls=, x=, y=, repeatCount=, xlink:href=,
  % properties
  margin:, padding:, background-image:, border:, top:, left:, position:, width:, height:, margin-top:, margin-bottom:, font-size:, line-height:,
  % CSS3 properties
  transform:, -moz-transform:, -webkit-transform:,
  animation:, -webkit-animation:,
  transition:,  transition-duration:, transition-property:, transition-timing-function:,
  }
}

@@

\begin{figure}[t]
  \begin{lstlisting}[style=htmlcssjs,language=JavaScript]
function safeAdd() {
  var iframe = document.createElement("iframe");
  var whereto = document.getElementById("debug");
  var frameid = document.getElementById("frame");
  iframe.setAttribute("id", "frame");
  if (frameid) { 
    iframe.removeChild(frameid.childNodes[0]); |c \label{lbl:error} |c
    // frameid.parentNode.removeChild(frameid); |c \label{lbl:correction} |c
  } else { whereto.appendChild(iframe); } }  
  \end{lstlisting}

  \begin{lstlisting}[style=htmlcssjs, language=HTML5]
<html>
  <body>
    <div id="frame"></div>
    <div id="debug"></div>
  </body>
</html>
  \end{lstlisting}

  \begin{lstlisting}[style=htmlcssjs, language=HTML5]
<html>
  <body>
    <div id="debug"></div>
  </body>
</html>
\end{lstlisting}    
  \caption{Stackoverflow Example}
  \label{fig..safeAdd.js}
\end{figure}

\begin{figure}[t]
  \begin{lstlisting}[style=htmlcssjs,language=JavaScript]
/*t dom */
function revealAll() {
 for(var row=0; row<2; row++){
  for(var col=0; col<2; col++){
   var obj = document.getElementById('square_'+row+'_'+col);
   var spans = obj.getElementsByTagName('SPAN');
   spans[0].style.display='';
   spans[1].style.display='none';
   spans[1].style.color='#000000'; } } }
  \end{lstlisting}

  \begin{lstlisting}[style=htmlcssjs, language=HTML5]
<html>
  <body>
    <div id="square_0_0">
      <span></span><span></span>
    </div>
    <div id="square_0_1">
      <span></span><span></span>
    </div>
    <div id="square_1_0">
      <span></span><span></span>
    </div>
    <div id="square_1_1">
      <span></span><span></span>
    </div>
  </body>
</html>
  \end{lstlisting}  
  \caption{Sudoku Example}
  \label{fig.revealAll.js}
\end{figure}

%% --------------------------------------------------------------------
\section{Test Generation Framework}
\label{sec.framework}
%% --------------------------------------------------------------------

@@

    \begin{call}{user}{fun.js + config}{server}{\textbf{end}}
      \begin{call}{server}{Initialize(fun.js)}{server}{(cfg, branches, cinfo, ifun, sig)}
      \end{call}

@@

Given JS function $f$ and some set of its branches $B_f$, the purpose of our testing framework is to generate input arguments for $f$ that trigger executions covering all branches in $B_f$. The general workflow of the test generation process is represented as a sequence diagram in Figure~\ref{fig.framework.architect}. The diagram consists of four interacting components: \User, \Server\, \Validator and \Client. \Server is the key one. It performs all preparatory steps (\emph{Initialization Phase}) as well as controls the data generation (\emph{Genetic Phase}). These two phases will be covered in details in Section~\ref{sub.sec.init.phase} and \ref{sub.sec.genetic.phase} respectively. Every request sent to \Validator\footnote{\url{https://validator.w3.org/nu/}} with an HTML document as the body triggers a response containing the list of syntactic errors in that document. \Client is a \emph{Node.js}\footnote{\url{https://nodejs.org/en/}} application that is mainly responsible for the execution of the JavaScript code. It relies on the \emph{jsdom}\footnote{\url{https://github.com/tmpvar/jsdom}} library simulating calls to native DOM.

Let us get a closer look at the sequence of interactions in Figure~\ref{fig.framework.architect}. First, \User provides JS file \emph{fun.js} that defines the function under test (FUT) \emph{fun} annotated with its type. Additionally, user supplies a configuration file \emph{config} that parameterizes some settings related to the genetic algorithm, random generation, logging etc. These data are processed by \Server in the \emph{Initialization Phase}. As an output we get a tuple with the following information about $fun$: control flow graph (\emph{cfg}), list of branches (\emph{branches}), static information about constants (\emph{cinfo}), instrumented function (\emph{ifun}) and the function's signature (\emph{sig}). \Server sends a \texttt{POST} request to the \Client to store \emph{ifun} and \emph{sig}. Second, \Server asks \User to select a branch(s) that has to be covered by the execution of \emph{fun}. For each branch our framework enters the \emph{Genetic Loop}. Inside of this phase it starts with a randomly generated population (\emph{Random Loop}), and continues to optimize it until earthier perfect fitness is found or generation limit is exceeded. Each candidate in the population has to be evaluated (\emph{Evaluation Loop}). To do so \Client parses the candidate according the known signature \emph{sig}, invokes \emph{ifun} on the result, collects some information along the execution and sends it back to \Server, where the final fitness value is computed for the candidate. New population is formed as the joint result of the \emph{Crossover} and \emph{Mutation Loop}. These two loops together with the \emph{Random} one are performed in conversation with \Validator. It forces to proceed generation until no more errors are present in the data. At the end of the \emph{Fitness Loop} the best candidate with the best fitness value are reported back to \User. After exploring all branches \emph{Genetic Loop} terminates.

@@

\begin{algorithm}[t!]
  \caption{Initialization Phase}
  \label{alg.init}
  \algsetup{linenosize=\tiny}
  \scriptsize
  \DontPrintSemicolon
  \SetAlgoVlined
  \SetKwInOut{Input}{Input}
  \SetKwInOut{Output}{Output}
  \SetKwInOut{Require}{Require}
  \SetKwProg{Fn}{Function}{}{}
  \Input{
    $fun.js$ is a JS file with FUT and type annotation
  }
  \Output{
    tuple with $(cfg, branches, cinfo, ifun, sig)$
  }
  \Fn{$Initialize(fun.js)$}{

    $(ast, sig) \longleftarrow ParseFunAndSig(fun.js)$\label{alg.init.parse.fun.sig}\;
    $cinfo \longleftarrow GetConstInfo(ast)$\label{alg.init.collect.info}\;
    $ast \longleftarrow NormalizeAST(ast)$\label{alg.init.transform.ast}\;
    $cfg \longleftarrow BuildCFG(ast)$\label{alg.init.build.cfg}\;
    $branches \longleftarrow GetBranches(cfg)$\label{alg.init.get.branches}\;
    $ifun \longleftarrow Instrument(ast)$\label{alg.init.instr}\;
    \Return{$(cfg, branches, cinfo, ifun, sig)$}
    }
\end{algorithm}

The goal of the \emph{Initialization Phase} is to prepare FUT for the follow up \emph{Genetic Phase}. The main steps of the initialization are illustrated in Algorithm~\ref{alg.init}. Given the file \emph{fun.js} as an input, we separately parse the function and its signature in line~\ref{alg.init.parse.fun.sig}. Currently our framework recognizes signatures composed of four primitive types such as \texttt{bool}, \texttt{int}, \texttt{float}, \texttt{string} and a newly introduced type \texttt{dom}. To facilitate random generation in the \emph{Genetic Phase} we statically analyze the FUT to collect useful data about potential string and numeric constants (line~\ref{alg.init.collect.info}). Moreover, we track calls to such browser API methods as \texttt{getElementsByTagName}, \texttt{getElementById}, \texttt{getElementsByClassName}, \texttt{getElementsByName}, which gives us information about \emph{tags}, \emph{ids}, \emph{classes} and \emph{names} respectively. Knowing this data significantly increases our chances to produce a desired test input. Sometimes it is necessary to normalize FUT before applying instrumentation (line~\ref{alg.init.transform.ast}), e.g. we transform \emph{if-then} branches to \emph{if-then-else}. After the normalization we can build CFG (line~\ref{alg.init.build.cfg}) and get all branches (line~\ref{alg.init.get.branches}). The final step is to instrument \emph{fun} (line~\ref{alg.init.instr}) in order to collect some run-time data such as execution trace, branch and loop distances, and constants. This data will be required for fitness evaluation (Section~\ref{sub.sub.sec.fitness.fun}) and updating of \emph{cinfo}.

@@

\begin{algorithm}[!t]
  \caption{Genetic Phase}
  \label{alg.gen}
  \algsetup{linenosize=\tiny}
  \scriptsize
  \DontPrintSemicolon
  \SetAlgoVlined
  \SetKwInOut{Input}{Input}
  \SetKwInOut{Output}{Output}
  \SetKwInOut{Require}{Require}
  \SetKwProg{Fn}{Function}{}{}
  \Input{
    $b_N$ is target branch, $cfg$ is control-flow graph\\
    $cinfo$ is constant info, $config$ is configuration file\\
    $ifun$ instrumented JS function, $sig$ type signature
  }
  \Output{
     element $p^*$ on which the best fitness $f^*$ is achieved  
  }
  \Fn{$Genetic(b_N, cfg, cinfo, config, ifun, sig)$}{
    $(size_P, size_A, size_G, rate_C, rate_M) \longleftarrow Read(config)$\label{alg.gen.read.config}\;
    $archive \longleftarrow \emptyset$; $i \longleftarrow 0 $\label{alg.gen.arch.init}\;
    $pop \longleftarrow Random(cinfo, size_P)$\label{alg.gen.random.gen}\;
    \Repeat{$isPerfect(f^*) \vee i > size_G $}{\label{alg.gen.loop.begin}
      $scoredPop \longleftarrow \emptyset$\;
      \ForAll{$p \in pop$}{
        $args \longleftarrow MakeArguments(p, sig)$\;
        $(tr, d_B, d_L, consts) \longleftarrow ifun(args)$\;
        $(f, p) \longleftarrow Evaluate(b_N, cfg, tr, d_B, d_L)$\;
        $scoredPop \longleftarrow (f, p) \cup scoredPop$\;
        $cinfo \longleftarrow consts \cup cinfo$\;
      }
      $combo \longleftarrow scoredPop \cup archive$\;
      $cross \longleftarrow Crossover(combo, rate_C)$\;
      $mut \longleftarrow Mutation(combo, cinfo, rate_M)$\;
      $pop \longleftarrow cross \cup mut$\;
      $archive \longleftarrow GetBest(combo, size_A)$\;
      $(f^*, p^*) \longleftarrow head(archive)$\;
      $i \longleftarrow i + 1$\;
    }\label{alg.gen.loop.end}
    \Return{$(f^*, p^*)$} 
  }
\end{algorithm}

The heart of our test data generation framework is in the \emph{Genetic Phase}. The main workflow of the underlying genetic algorithm (GA) is quite standard~\cite{poli2008field}, and with some adaptations presented in Algorithm~\ref{alg.gen}. We start with the reading of the \emph{config} file in line~\ref{alg.gen.read.config}. As a result we get information about the number of elements in a population ($size_P$), size of the archive ($size_A$), number of generations ($size_G$) and respectively crossover ($rate_C$) and mutation ($rate_M$) rates. \emph{Archive} has the size of  $size_A$ and stores the most fitted candidates observed so far. It is empty in the beginning of the evolution (line~\ref{alg.gen.arch.init}). First population of $size_P$ is randomly generated (line~\ref{alg.gen.random.gen}) based on the values passed through the \emph{cinfo}. Inside of the loop (lines~\ref{alg.gen.loop.begin}-\ref{alg.gen.loop.end}) we are trying optimize the ....

Then, this population is evaluated against given fitness function. If the perfectly fitted candidate is found or the budget is exhausted, the algorithm terminates and reports the best candidate so far. Otherwise it enters an iterative phase where after a selections procedure some candidates are crossed over and others are mutated. Afterwards a new population is formed and the evaluation repeats.

@@

\begin{description}
\item[\textbf{Q1}] How often \emph{validator} rejects generated HTML in case of the random and genetic (mutation and crossover) phases. This question characterizes potential performance issues that can occur in the generation.
\item[\textbf{Q2}] Efficiency (compare with random)  
\end{description}
